### 2022-01-19 papers/paper/preprint.tex
major facelift for the flowchart

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, fleqn, nicefrac}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}
\usepackage{fontawesome5}

@@

\begin{table}
    \centering
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name           & Symbol           & Conceptual framework                                                                 & Section               \\
        \midrule
        \Gls{slr}      & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
        Storm surge    & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
        Discount rate  & $1-\gamma$       & Fixed at 2\%                                                                         & \ref{sec:ead}         \\
        Depth-damage   & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
        Elevation cost & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
        Initial height & $h_0$            & Deterministic: equal to \gls{bfe} unless otherwise noted                             & TBD                   \\
        \bottomrule
    \end{tabular}
    \caption{Uncertainties modeled}\label{tab:uncertainties}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=4in]{xlrm.pdf}
    \caption{
        Conceptual diagram of the decision problem.
        A \gls{sow} consists of a description of the uncertain factors (red).
        We model a single decision (yellow).
        For each combination of \gls{sow} and policy, the system model (gray) is used to calculate metrics describing the performance of the policy over the selected \gls{sow}.
    }
\end{figure}

@@

Deeply uncertain sea level rise is the focus of our model.
Our analysis considers the common case where a decision maker or analyst has access to an existing set of simulations.
Specifically, we use experiments using the BRICK model version 0.2 \citep{wong_brick0.2:2017} simulations published in \citet{ruckart_hazardous:2008}.
There are eight \glspl{pdf}, each with ZXX\james{count} simulations: four emissions pathways (\gls{rcp} scenarios) $\times$ two models of ice sheet dynamics.

For consistency with literature on decision making under deep uncertainty \citep[see][]{marchau:2019}, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \acrfull{sow}}.

@@

\begin{figure}
    \centering
    \includegraphics[height=1.5in]{historic_surge.pdf}%
    \includegraphics[height=1.5in]{surge_return_period.png}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (a):
        time series of historic storms.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
        (b):
        return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull rule.
        Blue line shows a maximum likelihood GEV fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
    }
    \label{fig:observed-surges}
\end{figure}

@@

In \cref{fig:scenario-map}, we plot each possible future as a single dot for each of several possible amounts by which to heighten the house.
On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
On the $x$ axis is mean sea level in 2100, which is an interpretable 1-dimensional projection of the \glspl{sow}.
Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ.

@@

This plot sheds light on the question: ``how well does a given decision perform under different possible futures?''
The best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible.
However, the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
When $\Delta h$ increases, the construction cost increases but the variability between different \glspl{sow} decreases.
This plot suggests that $\Delta h = \SI{8}{ft}$ dominates $\Delta h = \SI{14}{ft}$ because \SI{8}{ft} is sufficient to make expected annual damages negligible at a substantially lower construction cost.

Yet important questions remain -- since the \glspl{sow} are not draws from the full \gls{pdf}

@@

\subsection{One size does not fit all}

\begin{enumerate}
    \item Tailoring guidance to specific building characteristics can improve outcomes
    \item Elevating to ``\gls{bfe} plus a foot'' is not always optimal \citep[this is consistent with][]{zarekarizi_suboptimal:2020,xian_elevation:2017}
    \item Both over- and under-building can be costly \citep{ansar_bigisfragile:2017,DossGollin:2019}
\end{enumerate}

\begin{figure}
    \centering
    \begin{subfigure}{0.48\textwidth}
        \subcaption{\SI{2.5}{ft} below BFE}
        \includegraphics[width=\textwidth]{5.5.pdf}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\textwidth}
        \subcaption{\SI{0.5}{ft} below BFE}
        \includegraphics[width=\textwidth]{7.5.pdf}
    \end{subfigure}
    \caption{
        Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \cite{wong_brick0.2:2017} dynamics).
        Note unequal $y$-axis scales.
    }
\end{figure}

\subsection{It matters which PDF you design for}

Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie, the \textbf{\itshape{Multiple PDF Problem}}.
Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
\begin{figure}
    \centering
    \includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:boxplots}
\end{figure}

@@

@@

While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability \citep[see][]{schneider_dangerous:2001,schneider_scenarios:2002}.
\begin{figure}
    \centering
    \includegraphics[width=4in]{bayes-rdm.pdf}
    \caption{
        ``Prior Model Averaging'' framework.
    }\label{fig:flowchart}
\end{figure}
Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
A system model $(f)$ quantifies the performance $(u)$ of a decision ($x$) under a single SOW ($s$).
To synthesize, we weight PDFs using an approach based on Bayesian Model Averaging \citep{wong_surge:2018,Yao:2018bu}.

@@

In this section we present a general framework for decision making under deep uncertainty.
We begin with terms and definitions that are broadly consistent with common adaptive frameworks\james{add refs}\klaus{did I read this right?}.
In \cref{sec:methods-weight} we present a Bayesian model for weighting \glspl{sow}.
In \cref{sec:methods-iterative} we discuss how this framework can be integrated into frameworks for iterative engagement with experts and stakeholders to expand and frame the model.

We start by modeling the performance of a decision $x \in \mathcal{X}$.
Outcomes depend both on the decision and on a state of the world $s \in \Omega$ that describes all relevant uncertainties.
To model outcomes $u \in \mathcal{U}$, we use a system model $f: \mathcal{X}, \Omega \rightarrow \mathcal{U}$ \citep[we generally follow the notation of][]{mcphail_robustness:2019} which we assume to be deterministic.
If $f$ is a stochastic function, we can define its metrics in terms of expectations, \eg $u_{ij} = \mathbb{E}\qty[u | x=x_i, s=s_j]$, which can be approximated through techniques like Monte Carlo sampling.

The literature on decision making under deep uncertainty \citep[see][for a review]{marchau:2019} emphasizes the importance of using system models like $f$ in an exploratory context by sampling widely from plausible futures $s \in \Omega$ and studying their implications \citep{bankes:1993,lempert_shaping:2003,kwakkel:2019,lamontagne_discovery:2018}.
This is often described as ``bottom-up'' modeling \citep{Brown:2012kb} and has parallels to model-based stress tests and vulnerability assessments.

Given our system model $f$ and a sampling of $J$ possible states of the world $s_1, s_2, \ldots, s_J$, we can calculate a set of \emph{\gls{sow} metrics} $u_{ij}$ that tell us ``how will decision $x_i$ perform under state of the world $s_j$'' for many $s_j$, and one or more $x_i$.
For an apples to apples comparison of how different decisions perform on the $j$th \gls{sow}, we can directly compare $u_{1j}, u_{2j}, \ldots, u_{Ij}$.
However, if we wish to compare the performance of two or more decisions across \emph{all} scenarios, we will need some way of aggregating the \gls{sow} metrics into \emph{decision metrics} $m_1, m_2, \ldots, m_I$ such that for every $x_i$ there is a single (scalar or vector) $m_i$.\footnote{I think the link back to our original motivation -- the multiple PDF problem}

\citet{mcphail_robustness:2019} split the problem of aggregating \gls{sow} metrics into decision metrics into two steps: ``subsetting'' (\eg, single value, all values, subset of values) and ``metric calculation'' (\eg, mean, variance, skew, kurtosis).\klaus{You say something about see discussion in \citet{mcinerney_robust:2012} -- please elaborate}
We generalize this to
\begin{equation}
    m_i = g(\qty{u_{i1}, u_{i2}, \ldots, u_{iJ}}, \qty{w_1, w_2, \ldots, w_J}),
\end{equation}
where $g$ is an \emph{aggregating function} describing the ``metric calculation'' step and the $w_j$ describe weights given to each \gls{sow}.
A optional, but likely useful, interpretation of the weighted samples $\qty{w_j, u_{ij} \forall j=1, \ldots, J}$ is an empirical \gls{pdf}, in which case $g$ maps a \gls{pdf} of future outcomes to the metrics $m_i$.\footnote{perhaps this belongs in the Discussion}
This is a flexible framework; \cref{tab:literature-weighting} shows some methods for robust decision making from the literature using our notation.\james{Add some references to the table}

\begin{table}
    \centering
    \footnotesize
    \begin{tabular}{llll}
        \toprule
        Citation        & Explanation                     & $w_j$                                                                             & $g(u, w)$          \\
        \midrule
        Many            & All $s_j$ are equally plausible & $w_j = \nicefrac{1}{J}$                                                           & any                \\
        Many            & Worst case                      & $w_j = \begin{cases}1 & u_{ij} = \min_{j'} u_{ij'} \\ 0 & \text{else}\end{cases}$ & $u$                \\
        TBD             & Average over \glspl{sow}        & $w_j = \nicefrac{1}{J}$                                                           & $\sum_j w_j u_{j}$ \\
        Characklis/Reed & 95th percentile                 & $w_j = \nicefrac{1}{J}$                                                           & weighted quantile  \\
        \bottomrule
    \end{tabular}
    \caption{
        A non-exhaustive summary of methods for setting $w_j$ and $g$ from the literature.
        (\emph{What other examples should we add?})
    }\label{tab:literature-weighting}
\end{table}

In the next section we propose an alternative scheme for weighting scenarios.

\paragraph{To think  about}
I have two flavors of this idea presented here!
\Cref{fig:framework}\james{Updates for \cref{fig:framework}: Don't need to say here that $f$ is deterministic. Larger text and darker color for questions. Rename $p(s)$ as ``prior over parameters and model structures''} takes a more Bayesian focus: the goal is to reason about properties of the \emph{distribution} over outcomes $p(u | x=x_i)$.
The text follows \citet{mcphail_robustness:2019} more.
I'm not sure which is easier to understand.

\subsection{Weighting \glsentryplural{sow}}\label{sec:methods-weight}

If the common practice of weighting all scenarios equally is applied ($w_j \propto 1$ in \cref{tab:literature-weighting}), a tension emerges: we want to \emph{explore} highly unlikely but theoretically possible trajectories to understand their consequences, but we don't want extremely unlikely scenarios to dominate our decision.

Some aims:
\begin{enumerate}
    \item Assumptions should be as transparent and visible as possible to facilitate critique and dialogue\klaus{I moved this up}
    \item We should be invariant to the parameterization of $s$ -- if we sample population uniformly from 1 million to 10 million or $\log_{10}$ population uniformly from \num6 to 7, we should get the same answer
    \item We should be invariant to the sampling of $s$:
          \begin{enumerate}
              \item if we include scientifically interesting but extremely low-probability scenarios (\eg, if someone created \gls{rcp} 45), our esimtates of $m_i$ should not change meaningfully.\james{Explain why}
              \item if we include nearly redundant scenarios (\eg, if we are studying \gls{rcp} 2.6, 4.5, 6.0, and 8.5 and we then add results from a \gls{rcp} 6.01 experiment), our estimates of $m_i$ should not change meaningfully.\james{Explain why}
          \end{enumerate}
\end{enumerate}
We can do this with probability if we have a target distribution $p(s)$.
We return to the question of how to choose this in \cref{sec:methods-prior}.

\subsection{Melding prior over \glsentryplural{sow}}\label{sec:methods-prior}

Often (including in the case study described in \cref{sec:case}) our \glspl{sow} are the outputs of complicated model chains.
To set our prior $p(s)$ we might like to incorporate prior external knowledge about both the

\subsection{Opportunities for iterative model expansion}\label{sec:methods-iterative}

\paragraph{Robustness checks}
The computationally expensive piece is computing the $u_{ij}$.
Once you do that, you can look at several $p(\Omega)$ and compute the $m_i$ multiple times.

\paragraph{Links to DMDU Methods}
You can use scenario discovery on the $(x_i, \mu_{ij}$).
(Others to note?)

\paragraph{Bayesian model checks}
Let's just cite \citet{gelman_workflow:2020} and \citet{gabry_visualization:2019} and point out that you can and should conduct sample-based checks on your prior $p(\Omega)$ and your posterior $p(u | x=x_i)$.\footnote{See discussion above about whether we want to emphasize that we are treating this as a posterior or not}

\section{Case study: house elevation}\label{sec:case}

We start with a bare-bones version of the \citet{zarekarizi_suboptimal:2020} model.
For interpretability, we fix the discount rate, house lifetime, and depth-damage curve to illustrate how our framework can be used to model nonstationary flood risk.
Real-world decision makers should consider additional uncertainties and stakeholder-identified metrics.

\begin{figure}
    \centering
    %\includegraphics[height=2.5in, width=0.6\textwidth, keepaspectratio=true]{xlrm}
    \caption{
        Diagram showing uncertainties ($X$; red), levers ($L$; yellow), system relationships ($R$; gray), and metrics ($M$; blue).
    }\label{fig:xlrm}
\end{figure}

\subsection{Scenarios}

Coastal floods are often decomposed into changes in local sea level \citep{kopp_evolving:2017,kopp_probabilistic:2014,wong_brick0.2:2017} plus storm surge \citep{garner_slrise:2018,cagigal_emulator:2020,rueda_surge:2016}.
We neglect waves and hydrodynamics for simplicity.

\gls{slr}:
\begin{enumerate}
    \item We have 2 models of ice sheet dynamics (3? Vivek and Tony are trying to tell me what models the data from \citet{ruckert_coastal:2019} actually comes from)
    \item We have 4 \gls{rcp} scenario
    \item We use the BRICK model \citep{wong_brick0.2:2017} simulations from \citet{ruckert_coastal:2019} to define $J$ \glspl{sow}.
\end{enumerate}
Storm surge:
\begin{enumerate}
    \item We have a $N$ year observational record
    \item We separate out storm surge from the background trend in some standard way
    \item We model storm surge as a stationary GEV model
    \item We place a prior on the return level (data space) rather than a prior on the parameters themselves (except the shape parameter -- we put a restrictive prior for it to be positive)\james{arguably we should stick to boring priors here -- the question being whether it's better to say we use iterative checks to improve the priors, or whether we want to keep things simple}
    \item We draw $K$ sets of parameters $\qty(\mu_k, \sigma_k, \xi_k)$
\end{enumerate}
Thus a \gls{sow} is defined just an 81 year time series of \gls{lsl}.
We will see what we do with the $K$ draws of storm surge parameters shortly.
\begin{figure}
    \caption{
        INSERT HERE: figure showing time series of observed surge and \gls{lsl} with notable storms indicated, as well as projections of future \gls{lsl} under all eight models (two BRICK, four \gls{rcp})
    }
\end{figure}

\subsection{Decisions}

We consider only one lever: how high to elevate the house today.
This is the same lever considered in \citet{zarekarizi_suboptimal:2020}

Three scenario metrics $u$:
\begin{enumerate}
    \item Construction cost
    \item Discounted expected flood damages
    \item Total cost (construction plus flood)
\end{enumerate}
The decision metrics $m$ are just the expected value of $u$.

\subsection{A simplified didactic system model}

\begin{enumerate}
    \item Costs of elevating the house are the same as \citet{zarekarizi_suboptimal:2020}
    \item Depth-damage: use HAZUS model from \citet{zarekarizi_suboptimal:2020}
    \item Discount: fix at some appropriate rate (2\%?)
    \item House lifetime: fix at some reasonable amount (80 years)
\end{enumerate}

Recall that we have $K$ simulations of the storm surge parameters.
For each year of each \gls{sow} we would like to estimate the expected value of the convolution of hazard (storm surge plus \gls{lsl}) with fragility (depth-damage).
A simple algorithm to estimate this is, for each \gls{pdf} of storm surge $k \in 1, \ldots, K$, to draw $N$ Monte Carlo sample values of storm surge, calculate the damage associated with each, and take the expectation with respect to $N$.
Then, take the expectation again with respect to the $K$ storm surge \glspl{pdf}.
Since this is computationally expensive, we build an emulator that estimates expected damage as a function of mean sea level only by fitting a spline to the result of systematically calculating expected damages for different values of \gls{lsl}.\footnote{A supplemental figure would be good here. This is probably a bit dense of a discussion.}

\begin{figure}
    \caption{
        INSERT HERE: figure showing the convolution of storm surge hazard through the depth-damage function.
        We fix \gls{lsl} constant but have many \glspl{pdf} of \gls{lsl}.
        This is the ``hazard, damage, risk'' graph that Klaus has sketched.
    }
\end{figure}

\subsection{Belief model}

\paragraph{Prior on outputs only}
Following \citet{garner_slrise:2018}, \citet{oddo_coastal:2017}, and \citet{lempert_slr:2012} we place a prior on \gls{slr} in 2100.\klaus{You suggest looking at another paper. Are you referring to this?\\\fullcite{sriver_sealevel:2018}}

\paragraph{Prior on inputs only}
Alternatively, we can place a prior over scenarios.
Let's say that we think that emissions forcing in 2100 is has some distribution.
This means we are placing a weight over \gls{rcp} scenarios.
We can assume both BRICK models are equally likely.

\paragraph{Melding}
We can do cool stuff with Bayesian melding \citep{poole_melding:2000,sevcikova_melding:2007}.
This combines the prior over inputs and outputs geometrically.
\begin{equation}\label{eq:melding}
    \text{melding equation here}
\end{equation}
This lets us express information about some inputs being more or less likely than others, and also some outputs being more or less likely than others.
THIS IS USEFUL because often we would like to incorporate relevant knowledge from many different sources on both inputs and outputs (\eg, IPCC reports, National Climate Assessments, \etc).

\paragraph{Implementation}
Code will be available on GitHub.
We use the Julia programming language.

\section{Results}\label{sec:results}

We will need a figure or two showing our prior and posterior checks, as well as some results.
Results should show the tradeoff between cost and damage under different assumptions.

\section{Discussion}

Most fundamentally, we are thinking about robustness a bit differently than usual.
The typical way that the DMDU literature thinks about it is robustness to a \gls{sow}.
We are thinking about robustness to different modeling structures and assumptions.
This has links to the ``rival framings'' literature that Pat has been working on.
Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the tradeoff curves as shown in figures XXXX (in \cref{sec:results}).

>>>>

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}
\usepackage{fontawesome5}
\usepackage[format=plain, textfont=it]{caption}

@@

\begin{table}
    \centering
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name           & Symbol           & Conceptual framework                                                                 & Section               \\
        \midrule
        \Gls{slr}      & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
        Storm surge    & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
        Discount rate  & $1-\gamma$       & Fixed at 2\%                                                                         & \ref{sec:ead}         \\
        Depth-damage   & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
        Elevation cost & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
        Initial height & $h_0$            & Deterministic: \SI{0.5}{ft} above \gls{bfe} unless otherwise noted                   & TBD                   \\
        \bottomrule
    \end{tabular}
    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=4in]{xlrm.pdf}
    \caption{
        Conceptual diagram of the decision problem.
        A \gls{sow} consists of a description of the uncertain factors (red).
        We model a single decision (yellow).
        For each combination of \gls{sow} and policy, the system model (gray) is used to calculate metrics describing the performance of the policy over the selected \gls{sow}.
    }\label{fig:xlrm}
\end{figure}

@@

Deeply uncertain sea level rise is the focus of our model.
Our analysis considers the common case where a decision maker or analyst has access to an existing set of simulations.
Specifically, we use simulations published in \citet{ruckart_hazardous:2008} that use the BRICK model version 0.2 \citep{wong_brick0.2:2017}.

\Cref{fig:boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways (\gls{rcp} scenarios) times two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    \includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:boxplots}
\end{figure}

For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \acrfull{sow}}.

@@

\begin{figure}
    \centering
    \includegraphics[height=1.5in]{historic_surge.pdf}%
    \includegraphics[height=1.5in]{surge_return_period.png}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (a):
        time series of historic storms.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
        (b):
        return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula .
        Blue line shows a maximum likelihood GEV fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
    }
    \label{fig:observed-surges}
\end{figure}

@@

The first that we consider is exploratory modeling.
In general, we consider $J$ simulated \glspl{sow} $\qty{s_1, \ldots, S_J}$ and $I$ possible decisions $\qty{x_1, \ldots, x_I}$.
As described in the previous section, these $J$ \glspl{sow} represent time series of mean sea level and the $I$ decisions represent how high to elevate the house.
For each combination of these, we calculate the vector of metrics described in \cref{sec:sow-metrics}: $u_{ij} = f(x_i, s_j)$ using the system model $f$, which includes storm surge (\cref{sec:storm-surge}), damage (\cref{sec:ead,sec:led}), and calculation of additional metrics (\cref{sec:sow-metrics}).
This framework is illustrated in \cref{fig:flowchart} inside the box labeled Section 3\klaus{Any suggestions for how I can draw these regions better?}; we return in \cref{sec:multiple-pdf,sec:synthesizing} to add additional steps to the model.

\begin{figure}
    \centering
    \includegraphics[width=4in]{bayes-rdm.pdf}
    \caption{
        Outline of the generally applicable decision theoretic framework developed.
        We begin in \cref{sec:multiple-simulation} by presenting results using exploratory modeling, \ie nodes  (a), (b), and (c).
        In \cref{sec:multiple-pdf} we discuss nodes (d) and (e) to understand scenario-conditional probability distributions.
        Finally in \cref{sec:synthesizing} we add nodes (f) and (g) to synthesize across multiple \glspl{pdf}.
    }\label{fig:flowchart}
\end{figure}

Analysis of these $u_{ij}$ sheds light on the question: ``how well does a given decision perform under different possible futures?''
\Cref{fig:scenario-map} shows each possible future as a single dot for each of several possible values for $\Delta h$.
On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
On the $x$ axis is mean sea level in 2100, which is an interpretable 1-dimensional projection of the \glspl{sow}.
Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ.

@@

\Cref{fig:scenario-map} shows that the best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible.
However, the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
As $\Delta h$ increases, the construction cost increases but the variability between different \glspl{sow} decreases, up to a point.
\emph{No decision dominates} any other, including $\Delta h = 0$, in the formal sense of being optimal over all \glspl{sow}.

@@

Whereas the analysis in \cref{sec:multiple-simulation} shed light on how decisions perform under different \glspl{sow}, the metrics calculated represent the performance of a given decision-\gls{sow} combination, and \emph{do not shed light on the performance of the decisions themselves}.

An alternative framework is to interpret the \glspl{sow} as draws from each of eight \glspl{pdf}.
This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as IID draws from the distribution of outcomes.

Some general notation is shown in \cref{fig:flowchart} (additional terms are added in \cref{sec:synthesizing}).
We have $K=8$ models of sea level rise, which we refer to as $\qty{\mathcal{M}_1, \ldots, \mathcal{M}_K}$.

This allows us to reason formally about the \gls{pdf} of outcomes for each candidate decision, conditional on a model.
We can use these \glspl{pdf} of outcomes to calculate metrics describing the performance of decisions.

@@

Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.

@@

While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability \citep[see][]{schneider_dangerous:2001,schneider_scenarios:2002}.

Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
A system model $(f)$ quantifies the performance $(u)$ of a decision ($x$) under a single SOW ($s$).
To synthesize, we weight PDFs using an approach based on Bayesian Model Averaging \citep{wong_surge:2018,Yao:2018bu}.

@@

Most fundamentally, we are thinking about robustness a bit differently than usual.
The typical way that the DMDU literature thinks about it is robustness to a \gls{sow}.
We are thinking about robustness to different modeling structures and assumptions.
This has links to the ``rival framings'' literature.
Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the tradeoff curves as shown in figures XXXX (in \cref{sec:results}).

### 2022-01-19 tikz/bayes-rdm.tex
major facelift for the flowchart

@@

    \node [block] (decisions) at (0, 3) {
        Candidate decisions\\
        $\qty{x_1, x_2, \ldots, x_I} \in \mathcal{X}$
    };
    \node [block] at (-4, 3) (models) {
        Multiple PDFs\\
        $\qty{M_1, \ldots, M_K} \in \mathcal{M}$
    };
    \node [block] at (-4, 0) (simulations) {
        Simulated SOWs\\
        $\qty{s_1, \ldots, s_{N_k}}_k$
    };
    \path [line] (models) -- (simulations);
    \node [function] (model) at (0, 0) {
        System model\\
        $u_{ij} = f(x_i, s_j)$
    };
    \path [line] (simulations) -- (model);
    \path [line] (decisions) -- (model);

    \node [derived] (combined) at (4, -3) {
        Conditional distributions\\
        over outcomes\\
        $\hat{p} \qty(u | x_i, M_k)$
    };
    \path [line] (model) -- (combined);

    \node [function] (prior) at (-4, -3) {
        Prior\\
        over SOWs\\
        $p\qty(s)$
    };
    \node [derived] (weights) at (0, -3) {
        Model weights\\
        $w_1, \ldots, w_K$
    };
    \path [line] (simulations) -- (weights);
    \path [line] (prior) -- (weights);

    \node [derived] (aggregated) at (4, -6) {
        Aggregated distribution\\
        over outcomes\\
        $\hat{p}(u | x_i)$
    };
    \path [line] (combined) -- (aggregated);
    \path [line] (weights) -- (aggregated);
\end{tikzpicture}

>>>>

\pgfdeclarelayer{bg}    % declare background
\pgfdeclarelayer{bg2}    % declare background
\pgfdeclarelayer{bg3}    % declare background
\pgfsetlayers{bg3, bg2, bg, main}  % order of layers (main = standard layer)

@@

    \node [block, label={[]135:{\footnotesize (d)}}] at (-4, 3) (models) {
            Multiple models\\
            $\qty{M_1, \ldots, M_K}$
        };
    \node [block, label={[]135:{\footnotesize (a)}}] (decisions) at (0, 3) {
            Candidate decisions\\
            $\qty{x_1, x_2, \ldots, x_I}$
        };
    \node [block, label={[]135:{\footnotesize (b)}}] at (-4, 0) (simulations) {
            Simulated futures\\
            $\qty{s_1, \ldots, s_J} \in \Omega$
        };
    \path [line] (models) -- (simulations);
    \node [function, label={[]45:{\footnotesize (c)}}] (model) at (0, 0) {
            System model\\
            $u_{ij} = f(x_i, s_j)$
        };
    \path [line] (simulations) -- (model);
    \path [line] (decisions) -- (model);

    \node [function, label={[]135:{\footnotesize (f)}}] (prior) at (-4, -3.5) {
            Prior over\\
            possible futures\\
            $p\qty(\Omega)$
        };
    \node [derived, label={[]90:{\footnotesize (g)}}] (weights) at (0, -3.5) {
            Weights\\
            $w_1, \ldots, w_J$
        };
    \path [line] (simulations) -- (weights);
    \path [line] (prior) -- (weights);

    \node [derived, label={[]135:{\footnotesize (e)}}] (combined) at (4, 0) {
            Conditional\\
            distribution\\
            over outcomes\\
            $\hat{p}(u | x_i, M_k)$
        };
    \path [line] (model) -- (combined);
    \path [line] (weights) -- (combined);

    \node [rotate=45, color=thirds] (s3) at (-1.25, 1.25) {\large \textbf{Section 3}};
    \begin{pgfonlayer}{bg}    % select background
        \path [dotted, thick, color=thirds, fill=grigio!20] (-6, -1.5) -- (2, -1.5) -- (2, 4.25) -- (-2, 4.25) -- (-2, 1.25) -- (-6, 1.25) -- (-6, -1.5);
    \end{pgfonlayer}

    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\large \textbf{Section 4}};
    \begin{pgfonlayer}{bg2}    % select background
        \path [dotted, thick, color=thirds, fill=rosso!20] (-6.5, 4.6) -- (-6.5, -1.75) -- (5.5, -1.75) -- (5.5, 4.6) -- (-6.5, 4.5);
    \end{pgfonlayer}

    \node [color=thirds] (s5) at (4, -3.5) {\large \textbf{Section 5}};
    \begin{pgfonlayer}{bg3}    % select background
        \path [dotted, thick, color=thirds, fill=giallo!20] (-6.75, 4.8) -- (-6.75, -5) -- (5.75, -5) -- (5.75, 4.8) -- (-6.75, 4.8);
    \end{pgfonlayer}

\end{tikzpicture}
